Introduction_outline

### 2019-01-21 Methods.Rmd
Updated commit

setwd("~/Documents/MastersProject")

@@

upwelling_SACTN_daily_wc <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
upwelling_SACTN_daily_wc <- upwelling_SACTN_daily_wc[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell

SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_SACTN_daily_wc$index)

SACTN_daily_temps_indexsplit <- SACTN_daily_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  dplyr::rename(insitu_temp = temp)
```

@@

```{r}
SACTN_daily_temps_monthly <- SACTN_daily_temps %>% 
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"),
                                              "Spring","Error")))))
```

@@

Here I load the AVHRR CSV dataset. Large dataset and slows down the laptop. Here i split the data into xaa and xab. Here xaa and xab only include temperature collected for two years.
```{r}
# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")

@@

#OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
##################
# xaa <- read_csv("Data/xaa.csv",  col_names = c("lon", "lat", "temp", "date"))
# xab <- read_csv("Data/xab.csv",  col_names = c("lon", "lat", "temp", "date"))
# combined_OISST <- rbind(xaa, xab)
# save(combined_OISST, file = "data/combined_OISST.RData")
```

@@

# Select nearest 1 pixels (k = 1)
# here we use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(upwelling_SACTN_daily_wc[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = upwelling_SACTN_daily_wc$site)
```

@@

```{r}
matched_SST <- OISST_sites %>% 
  left_join(SACTN_daily_temps_indexsplit, by = c("site", "date")) %>% 
    na.trim()
```

```{r}
# MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
# MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
# MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
# MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
# MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
# MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
# MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
# MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
# MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
# 
# MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
#                  MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>% 
#   dplyr::rename(site = station)
# save(MUR_SST, file = "Data/MUR_SST.RData")
```

hee1 <- try1 %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = TRUE)) %>% 
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, temp, date)

ggplot(data = thee1, aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw() +
  theme(strip.text.x = element_text(size = 3))

>>>>

@@

upwelling_sites <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
upwelling_sites <- upwelling_sites[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell

SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_sites$index)

# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")

SACTN_daily_temps_indexsplit <- SACTN_daily_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  dplyr::rename(insitu_temp = temp)

save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
```

@@

```{r}
SACTN_monthly <- function(df) {
  SACTN_split_monthly <- df %>%
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(SACTN_split_monthly)
}

# SACTN_daily_temps_indexsplit <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
# save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
```

Find where different sites have overlapping time series within this `SACTN_daily_temps_indexsplit` dataset. A time series with a length of 3years is long enough.

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/SACTN_daily_temps_indexsplit.RData")
SACTN_3yr <- SACTN_daily_temps_indexsplit %>% 
  filter(year(date) %in% seq(1998, 1999, 2000))
```

```{r}
ggplot(data = SACTN_3yr , aes(x = date, y = insitu_temp)) +
  geom_hline(aes(yintercept = mean(insitu_temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
```

@@

```{r}
MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
                 MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))

save(MUR_SST, file = "Data/MUR_SST.RData")
```

```{r}
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp = nearest1)

match_func <- function(df){
  match <- SACTN_daily_temps_indexsplit  %>% 
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR <- match_func(df = MUR_SST)
```

```{r}
match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
  return(plot1)
}

MUR_plot <- match_plot(df = MUR_SST)
```

################################

-  Need to plot a line graph with all SST dat and the insitu data on one graph

################################

Here I load the AVHRR CSV dataset.

```{r}
# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")

@@

#OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
```

@@

# Select nearest 1 pixels (k = 1)
# here we use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(upwelling_sites[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = upwelling_sites$site)
```

@@

```{r}
insitu_OISST <- match_func(df = OISST_sites)
```

```{r}
OISST_plot <- match_plot(df = OISST_sites)
```

### 2019-01-21 Write_up/Writing/Introduction_outline.Rmd
Updated commit

---
title: "Introduction_outline"
author: "Amieroh Abrahams"
date: "13 January 2019"
output: html_document
---

@@

@@

## Extreme events

>>>>

---
title: "Introduction_outline"
author: "Amieroh Abrahams"
date: "13 January 2019"
output:
  pdf_document: default
  html_document: default
---

@@

### In situ data

Having coastal, in situ seawater temperature data is a rare privilege to researchers. Most biological and ecological research requires coastal temperature data. The South African Coastal Temperature Network (SACTN). The availability of the in situ collected coastal temperature data products for several decades and provides a reliable source of accurate coastal seawater temperature data (Smit et al. 2013). SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments. At present, this dataset contained temperature of 129 sites along the three distinct coasts of the South African coastline. Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneve per site. With the longest time series in the dataset being that of Gordons Bay, recorded by SAWS. Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily. During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.

The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors. For instance, these may be as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data. One problematic outcome of this sampling methodology is that due to specific work of the South African Weather Service (SAWS) and Kwazulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and records at a data precision of 0.500 ◦ C. However, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C (Jarraud, 2008). Because of these inconsistencies, this dataset required verification from an independent source before its use within research. Smit et al. (2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results. Using these in situ collected temperature data along with remotely sensed SST data requires a clear understanding in the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long term measurements are for use in climate change studies.

It must also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements (Aguilar et al., 2003). Specifically uncertainty lies in the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period. We are however certain that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs. Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.

For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset. Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers. These data points were then changed to NA (not available) so as to not interfere with analysis. All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/). The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MASTERSPROJECT.

@@

### Extreme events

Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells (e.g. Gunter, 1941; Lirman et al., 2011; Boucek et al., 2016) or heatwaves (e.g. Gordon et al., 1988; Stott et al., 2004; Perkins-Kirkpatrick et al., 2016). These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place. These events are often undetected but studies have found that extreme events have occurred as recently as approximately four decades ago. Jentsch et al. (2007) were the first to discuss this mode of climate change and brought it to focus. Steps have since been taken towards understanding the drivers, impacts and consequences of these events. Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming (REFERENCE). Extreme events are often associated with toxic algal blooms, shifts in species distribution, and the bleaching of coral reefs (McCabe et al., 2016; Harley and Paine, 2009; Schoepf et al., 2015).

Little is known about extreme events occurring within the South African region. Very little South African researchers spend their time and careers on studyig extreme events within this region. During the 1980s, a study done by Birkett and Cook was the only study done where they anaylsed the impact of extreme warming events on the life cycle of white muscles,Donax serra Röding. However, more recently studies done by Schlegel et al, (2017,2018) examined the relationship between the nearshore and offshore occurances of marine heatwaves and cold-spells as well as the influence of predominant atmosphetic and oceanic patterns during Coastal Marine Heatwaves. It is now becoming accepted that extreme events are a threat to several ecosystem (Thompson et al., 2013).
